---
title: "Energy_HolidayEffects"
output: html_document
date: "2022-11-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(here)
library(ggplot2)
library(RColorBrewer)
library(lubridate)
library(data.table)
library(dplyr)
devtools::load_all()
.d <- `[`
```


## Energy Data - Holiday Effects

This document should briefly motivate the inclusion of public holiday effects for predicting Energy Data.

```{r read data}
energy <- format.energy.byhour("20221109") |>
  .d(, month := month(utchour)) |>
  .d(, hour := hour(utchour)) |>
  .d(, weekday := wday(utchour))
```

For example, for October 3, the day of Germany Reunification:

```{r oct3, message = FALSE, warning = FALSE}
#get first five days of October for all years
holen <- energy |>
  filter(lubridate::month(utchour) == 10, lubridate::day(utchour) %in% seq(1:5))

#make factor that codes normal day - weekend day - public holiday
holen |>
  .d(, wend := as.numeric(weekday %in% c(1,7))) |> #get weekend days
  .d(lubridate::day(utchour) == 3, wend := 2) |> #get october 3
  .d(, wend := factor(wend, levels = c(0,1,2), 
                      labels = c("normal day", "weekend", "holiday"))) |>
  .d(, date := date(utchour)) #for plotting
```

```{r plotdat}
ggplot(holen, aes(x = hour, y = dhr, color = wend, group = date)) +
  geom_line() +
  scale_color_brewer(palette = "Set2") +
  theme_minimal()+
  ylim(30,80)

```

And for May 1 (same code as before)

```{r may1, message = FALSE, warning = FALSE, echo = FALSE}
#get first five days of October for all years
holen <- energy |>
  filter(lubridate::month(utchour) == 5, lubridate::day(utchour) %in% seq(1:5))

#make factor that codes normal day - weekend day - public holiday
holen |>
  .d(, wend := as.numeric(weekday %in% c(1,7))) |> #get weekend days
  .d(lubridate::day(utchour) == 1, wend := 2) |> #get october 3
  .d(, wend := factor(wend, levels = c(0,1,2), 
                      labels = c("normal day", "weekend", "holiday"))) |>
  .d(, date := date(utchour)) #for plotting
```

```{r plotdatmay1, echo = FALSE}
ggplot(holen, aes(x = hour, y = dhr, color = wend, group = date)) +
  geom_line() +
  scale_color_brewer(palette = "Set2") +
  theme_minimal()+
  ylim(30,80)

```

And for Christmas (classify 24th, 25th and 26th as holiday)
```{r christmas, message = FALSE, warning = FALSE, echo = FALSE}
#get first five days of October for all years
holen <- energy |>
  filter(lubridate::month(utchour) == 12, lubridate::day(utchour) %in% seq(22,27))

#make factor that codes normal day - weekend day - public holiday
holen |>
  .d(, wend := as.numeric(weekday %in% c(1,7))) |> #get weekend days
  .d(lubridate::day(utchour) %in% c(24,25,26), wend := 2) |> #get october 3
  .d(, wend := factor(wend, levels = c(0,1,2), 
                      labels = c("normal day", "weekend", "holiday"))) |>
  .d(, date := date(utchour)) #for plotting
```

```{r plotchristmas, echo = FALSE}
ggplot(holen, aes(x = hour, y = dhr, color = wend, group = date)) +
  geom_line() +
  scale_color_brewer(palette = "Set2") +
  theme_minimal() +
  ylim(30,80)

```

As we can see, the energy demand during public holidays is much more in line with that of weekends than that of normal weekdays. Christmas is a bit of an exception, with weekdays not separating as nicely from weekend/holiday days (likely due to this time of the year being a sort of holiday period aside from the two actual Christmas holidays plus Christmas Eve); nevertheless, these three holiday days are especially low in demand, so still makes sense to code them. 

All in all, it would make sense to thus include a public holiday effect.
